---
title: "Predict likelihood of Default"
author: "XW, PC"
date: "March 25, 2018"
output: 
  html_document:
    fig_width: 7
    fig_height: 4
    fig_caption: true
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list=ls())
df = read.csv("UCI_Credit_Card.csv")
colnames(df)[7] <- "PAY_1"
df <- df[!is.na(df$default.payment.next.month), ]
df$default.payment.next.month = as.factor(df$default.payment.next.month)
df <- subset(df,df$SEX!=0)
df <- subset(df,df$EDUCATION!=0)
df <- subset(df,df$MARRIAGE!=0)
df$EDUCATION[df$EDUCATION==6] <- 5
df$ID <- NULL
summary(df)
sapply(df,class)
```
Blindly predicts no default, accuracy = 23301/(23301+6631) = 77.87%

```{r}
barplot(table(df$SEX), main='SEX Bar Plot', xlab='Sex',names.arg = c("Male","Female"))
barplot(table(df$EDUCATION), main='Education Bar Plot', xlab='Education Level',
        names.arg = c("Graduate School","University","High School","Others","Unknown"))
barplot(table(df$MARRIAGE), main='Marriage Bar Plot', xlab='Marital Status',
        names.arg = c("Married","Single","Others"))
hist(df$AGE, xlab="Age", main='Histogram of Age')
```

```{r, warnings=FALSE, message=FALSE}
attach(df)
library(ggplot2)
sex=factor(SEX)
levels(sex)=c("Male","Female")
marriage=factor(MARRIAGE)
levels(marriage)=c("Married","Single","Others")
education=factor(EDUCATION)
levels(education)=c("Graduate School","University","High School","Others","Unknown")

ggplot(df,aes(x=sex,fill = default.payment.next.month))+geom_bar()+
  ggtitle("Gender Profile of Default Payment Vs Non-Default")+
  xlab("Gender")+
  theme(legend.position="top")

ggplot(df,aes(x=marriage,fill = default.payment.next.month))+geom_bar()+
  ggtitle("Marital Status Profile of Default Payment Vs Non-Default")+
  xlab("Marital Status")+
  theme(legend.position="top")

ggplot(df, aes(x = education, fill = default.payment.next.month)) + geom_bar() +
  ggtitle("Education Profile of Default Payment Vs Non-Default")+
  xlab("Education") +
  theme(legend.position="top")

# Density plots 
ggplot(df,aes(x=AGE)) + geom_density(aes(color = default.payment.next.month ))+
  ggtitle("Age Profile of Default Payment Vs Non-Default")+
  ylab("Denisty")+
  theme(legend.position="top")


ggplot(df,aes(x=marriage,y=LIMIT_BAL/1000,fill=sex))+geom_boxplot()+
  xlab("Marital Status")+
  ylab("Balance Limit")+
  theme(legend.position="top")

plot(factor(education),LIMIT_BAL/1000,main='Balance Limit vs. Education', xlab='Education Level', ylab="Balance Limit")

ggplot(df,aes(x=default.payment.next.month),aes(y=stat_count((SEX))))+geom_bar(aes(fill=factor(education)))+
  xlab("Default Status")+
  ylab("Count")+
  facet_wrap(~ factor(sex))+
  theme(legend.position="top")
```

```{r,warnings=FALSE,message=FALSE}
library(dplyr)
##Calculating monthly default rate from April 2005 to September 2005
default_rate=vector(length=6)
names(default_rate)= c("04-2005","05-2005","06-2005","07-2005","08-2005","09-2005")
default_rate[6]= (nrow(filter(df,PAY_1 > 2 ))/nrow(df))*100
default_rate[5]= (nrow(filter(df,PAY_2 > 2 ))/nrow(df))*100
default_rate[4]= (nrow(filter(df,PAY_3 > 2 ))/nrow(df))*100
default_rate[3]= (nrow(filter(df,PAY_4 > 2 ))/nrow(df))*100
default_rate[2]= (nrow(filter(df,PAY_5 > 2 ))/nrow(df))*100
default_rate[1]= (nrow(filter(df,PAY_6 > 2 ))/nrow(df))*100

months_name=names(default_rate)
# Default(rate)=default > 60 days vs months
data_rate = data.frame(months=months_name, default_rate=default_rate)
ggplot(data=data_rate,aes(x=months, y=default_rate))+geom_point()+
  ggtitle("Default rate(default payment > 2 months) Vs. Months")+
  xlab("Months")+
  ylab("Default Rate (%)")+
  geom_point()+
  geom_line()
```
Default rate is linearly increasing from April 2005 to August 2005 and then slowing down in September 2005.


```{r, warnings=FALSE,message=FALSE}
library(corrplot)
cor=cor(subset(df,select=c(LIMIT_BAL,SEX,EDUCATION,MARRIAGE,AGE,PAY_1,PAY_2,PAY_3,PAY_4,PAY_5,PAY_6,
                           BILL_AMT1,BILL_AMT2,BILL_AMT3,BILL_AMT4,BILL_AMT5,BILL_AMT6,
                           PAY_AMT1,PAY_AMT2,PAY_AMT3,PAY_AMT4,PAY_AMT5,PAY_AMT6)))
corrplot(cor,method="circle")

dcor=cor(subset(df,select=c(LIMIT_BAL,
                           BILL_AMT1,BILL_AMT2,BILL_AMT3,BILL_AMT4,BILL_AMT5,BILL_AMT6,
                           PAY_AMT1,PAY_AMT2,PAY_AMT3,PAY_AMT4,PAY_AMT5,PAY_AMT6)))
corrplot(dcor,method="circle")

```
The above correlation plot shows a strong correlation between billing amounts in each month and no correlation between payment status in each month.

High correlation between PAY, BILL_AMT
one in BILL_AMT


```{r}
set.seed(1)
train_ind = sample(1:nrow(df),0.75*nrow(df))
df_train = df[train_ind,]
df_test = df[-train_ind,]
df_logit = glm(default.payment.next.month ~ ., data = df_train, family = binomial)
summary(df_logit)
```
LIMIT_BAL, SEX, EDUCATION, MARRIAGE, AGE, PAY_1, PAY_2, PAY_3, BILL_AMT1, PAY_AMT1, PAY_AMT2

```{r,warnings=FALSE,message=FALSE}
library(ROCR)
logit_pred = predict(df_logit, df_test, type = "response")
pred <- prediction(logit_pred, df_test$default.payment.next.month)
perf <- performance(pred,"tpr","fpr")

# plotting the ROC curve
plot(perf,col="black",lty=3, lwd=3)
auc <- performance(pred, measure = "auc")
auc <- auc@y.values[[1]]
auc
# KS STATISTIC
ks <- max(perf@y.values[[1]]-perf@x.values[[1]])
ks


library(MLmetrics)
#KS_Stat(y_pred = logit_pred, y_true = df_test$default.payment.next.month)
default_num=as.numeric(df_test$default.payment.next.month)
Gini(y_pred = logit_pred, y_true = default_num)

library(pROC)
my_roc <- roc(df_test$default.payment.next.month, logit_pred)
coords(my_roc, "best", ret = "threshold")
logit_bin = as.numeric(logit_pred >=coords(my_roc, "best", ret = "threshold"))
table(logit_bin, df_test$default.payment.next.month)
output = table(logit_bin, df_test$default.payment.next.month)
acc1 = (output[1,1]+output[2,2])/sum(output)
acc1
```


```{r,warnings=FALSE,message=FALSE}
library(leaps)
regfit.full <- regsubsets(default.payment.next.month~.,data=df,nvmax=23)
reg.summary <- summary(regfit.full)
coef(regfit.full,9)
reg.summary$rss
reg.summary$rsq
which.min(reg.summary$rss) 
which.max(reg.summary$rsq)
```

```{r}
plot(reg.summary$adjr2,xlab="Number of Variables",ylab="Adjusted R Square", type="l")
which.max(reg.summary$adjr2)
points(15,reg.summary$adjr2[15],col="red",cex=2,pch=20)

plot(reg.summary$cp,xlab="Number of Variables",ylab="Cp", type="l")
which.min(reg.summary$cp)
points(15,reg.summary$cp[15],col="red",cex=2,pch=20)

plot(reg.summary$bic,xlab="Number of Variables",ylab="BIC", type="l")
which.min(reg.summary$bic)
points(9,reg.summary$bic[9],col="red",cex=2,pch=20)
```
9 variables, easy to interpret
9 -- LIMIT_BAL + EDUCATION + MARRIAGE + AGE + PAY_1 + PAY_2 + PAY_3 + BILL_AMT1 + PAY_AMT1
10 -- LIMIT_BAL + SEX + EDUCATION + MARRIAGE + AGE + PAY_1 + PAY_2 + PAY_3 + BILL_AMT1 + PAY_AMT1

```{r}
keep=c("LIMIT_BAL", "EDUCATION", "MARRIAGE", "AGE", "PAY_1", "PAY_2", "PAY_3", "BILL_AMT1", "PAY_AMT1","default.payment.next.month")
df_new=df[keep]
set.seed(1)
train_ind = sample(1:nrow(df_new),0.75*nrow(df_new))
df_train = df_new[train_ind,]
df_test = df_new[-train_ind,]
summary(df_train$default.payment.next.month)
df_logit = glm(default.payment.next.month ~., data = df_train, family = binomial)
summary(df_logit)
```

```{r message=FALSE, warning=FALSE}
library(ROCR)
logit_pred = predict(df_logit, df_test, type = "response")
pred <- prediction(logit_pred, df_test$default.payment.next.month)
perf <- performance(pred,"tpr","fpr")

# plotting the ROC curve
plot(perf,col="black",lty=3, lwd=3)
auc <- performance(pred, measure = "auc")
auc <- auc@y.values[[1]]
auc
# KS STATISTIC
ks <- max(perf@y.values[[1]]-perf@x.values[[1]])
ks
```

```{r message=FALSE, warning=FALSE}
library(pROC)
my_roc <- roc(df_test$default.payment.next.month, logit_pred)
coords(my_roc, "best", ret = "threshold")
logit_bin = as.numeric(logit_pred >=coords(my_roc, "best", ret = "threshold"))
table(logit_bin, df_test$default.payment.next.month)
output = table(logit_bin, df_test$default.payment.next.month)
acc1 = (output[1,1]+output[2,2])/sum(output)
acc1
```
```{r message=FALSE, warning=FALSE}
library(MLmetrics)
default_num=as.numeric(df_test$default.payment.next.month)
Gini(y_pred = logit_pred, y_true = default_num)
```

```{r}
set.seed(1)
train_ind = sample(1:nrow(df),0.75*nrow(df))
df_normal = scale(df[, !names(df) %in% 'default.payment.next.month'])
df_normal_train = df_normal[train_ind,]
df_normal_test = df_normal[-train_ind,]
train.default = df_new$default.payment.next.month[train_ind]
test.default = df_new$default.payment.next.month[-train_ind]
```

```{r}
#KNN
library(class)
k = c(1, 5, 10, 20, 50, 75, 100)
accuracy = rep(NA, length(k)) 
for (i in seq(length(k))){
  set.seed(1)
  knn.pred = knn(df_normal_train, df_normal_test, train.default, k=k[i]);
  accuracy[i] = mean(knn.pred == test.default)
}
accuracy
plot(k, accuracy, type = "b", col="red")
```

```{r}
#set.seed(1)
#knn.pred = knn(df_normal_train, df_normal_test, train.default, k=50)
#knn.pred1 <- attr(knn.pred, "prob")
pred <- prediction(knn.pred1, test.default)
perf <- performance(pred,"tpr","fpr")

# plotting the ROC curve
plot(perf,col="black",lty=3, lwd=3)
auc <- performance(pred, measure = "auc")
auc <- auc@y.values[[1]]
auc
# KS STATISTIC
ks <- max(perf@y.values[[1]]-perf@x.values[[1]])
ks
```

```{r}
x=model.matrix(default.payment.next.month~.,data.frame(df_normal))[,-1]
y=df$default.payment.next.month
```

```{r,warnings=FALSE, message=FALSE}
library(glmnet)
# Model fit on the training set
ridge.mod <- glmnet(x[train_ind,], y[train_ind], alpha=0, thresh=1e-12, family = "binomial")
# Use 10-fold cross-validation to choose lambda
set.seed(1)
cv.out=cv.glmnet(x[train_ind,], y[train_ind], nfolds = 10, lambda = ridge.mod$lambda, alpha=0, family = "binomial",type.measure = "mse")
plot(cv.out)
bestlam = cv.out$lambda.min
bestlam
```

```{r}
plot(ridge.mod,xvar="lambda")
abline(v=log(bestlam),col="red")
coef(cv.out,s=bestlam)
```

```{r,message=FALSE, warning=FALSE}
ridge.prob = predict(ridge.mod, s = bestlam, newx=x[-train_ind,], type="response")

pred <- prediction(ridge.prob, y[-train_ind])
perf <- performance(pred,"tpr","fpr")

# plotting the ROC curve
plot(perf,col="black",lty=3, lwd=3)
auc <- performance(pred, measure = "auc")
auc <- auc@y.values[[1]]
auc
# KS STATISTIC
ks <- max(perf@y.values[[1]]-perf@x.values[[1]])
ks
default_num=as.numeric(y[-train_ind])
Gini(y_pred = ridge.prob, y_true = default_num)

my_roc <- roc(y[-train_ind], ridge.prob)
coords(my_roc, "best", ret = "threshold")
ridge_bin = as.numeric(ridge.prob >=coords(my_roc, "best", ret = "threshold"))
table(ridge_bin, y[-train_ind])
output = table(ridge_bin, y[-train_ind])
acc2 = (output[1,1]+output[2,2])/sum(output)
acc2

```

```{r}
# Model fit on the training set
lasso.mod <- glmnet(x[train_ind,], y[train_ind], alpha=1, thresh=1e-12, family = "binomial")
# Use 10-fold cross-validation to choose lambda
set.seed(1)
cv.out=cv.glmnet(x[train_ind,], y[train_ind], nfolds = 10, lambda = lasso.mod$lambda, alpha=1, family = "binomial", type.measure = "auc")
plot(cv.out)
bestlam = cv.out$lambda.min
bestlam = exp(-3.4)
bestlam
```

```{r}
plot(lasso.mod,xvar="lambda")
abline(v=log(bestlam),col="red")
coef(cv.out,s=bestlam)
```


```{r message=FALSE, warning=FALSE}
lasso.prob = predict(lasso.mod, s = bestlam, newx=x[-train_ind,], type="response")
pred <- prediction(lasso.prob, y[-train_ind])
perf <- performance(pred,"tpr","fpr")

# plotting the ROC curve
plot(perf,col="black",lty=3, lwd=3)
auc <- performance(pred, measure = "auc")
auc <- auc@y.values[[1]]
auc

# KS STATISTIC
ks <- max(perf@y.values[[1]]-perf@x.values[[1]])
ks

Gini(y_pred = lasso.prob, y_true = default_num)

my_roc <- roc(y[-train_ind], lasso.prob)
coords(my_roc, "best", ret = "threshold")
lasso_bin = as.numeric(lasso.prob >=coords(my_roc, "best", ret = "threshold"))
table(lasso_bin, y[-train_ind])
output = table(ridge_bin, y[-train_ind])
acc3 = (output[1,1]+output[2,2])/sum(output)
acc3
```



```{r, warnings=FALSE, message=FALSE}
library(tree)
library(rpart)
library(rpart.plot)
fit <- rpart(default.payment.next.month ~ ., method="class", data=df, subset = train_ind, cp=0.001)
print(fit) # display the results 
par(mfrow = c(1,1), xpd = NA)
summary(fit)

fit$cptable

plotcp(fit)

rpart.plot(fit, tweak=1, type = 2, fallen.leaves = TRUE, main="Decision Tree") #fit.pruned
fit.pred <- predict(fit, df_test, type="class")
table(fit.pred,df_test$default.payment.next.month)
mean(fit.pred == df_test$default.payment.next.month)
caret::varImp(fit, surrogates = FALSE, competes = TRUE)

fit.pruned <- prune(fit, cp = 0.0016)
rpart.plot(fit.pruned, tweak=1, type = 2, fallen.leaves = TRUE, main="Decision Tree") 
fit.pred <- predict(fit.pruned, df_test, type="class")
table(fit.pred,df_test$default.payment.next.month)
mean(fit.pred == df_test$default.payment.next.month)
caret::varImp(fit.pruned, surrogates = FALSE, competes = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tree)
set.seed(1)
train_ind = sample(1:nrow(df),0.75*nrow(df))
df_train = df[train_ind,]
df_test = df[-train_ind,]
tree.default=tree(default.payment.next.month ~., df_train)
summary(tree.default)
tree.pred=predict(tree.default,df_test,type="class")
table(tree.pred,df_test$default.payment.next.month)
plot(tree.default)
text(tree.default,pretty=0)
set.seed(1)
cv.default=cv.tree(tree.default,FUN=prune.misclass)
cv.default
plot(cv.default$size ,cv.default$dev ,type="b")
plot(cv.default$k ,cv.default$dev ,type="b")
prune.default=prune.misclass(tree.default,best=2)
plot(prune.default)
text(prune.default,pretty=0)
prune.pred=predict(prune.default,df_test,type="class")
table(prune.pred,df_test$default.payment.next.month)
```

```{r}
tree.pred=predict(tree.default,df_test,type="vector")
pred <- prediction(tree.pred[,2], df_test$default.payment.next.month)
perf <- performance(pred,"tpr","fpr")

# plotting the ROC curve
plot(perf,col="black",lty=3, lwd=3)
auc <- performance(pred, measure = "auc")
auc <- auc@y.values[[1]]
auc
# KS STATISTIC
ks <- max(perf@y.values[[1]]-perf@x.values[[1]])
ks

library(MLmetrics)
default_num=as.numeric(df_test$default.payment.next.month)
Gini(y_pred = tree.default, y_true = default_num)
```

```{r,warnings=FALSE,message=FALSE}
library(randomForest)
#bagging
set.seed(1)
bag.default=randomForest(default.payment.next.month~.,data=df,subset=train_ind,mtry=23,ntree=200)
yhat.bag = predict(bag.default,newdata=df_test,type="class")
table(yhat.bag,df_test$default.payment.next.month)
mean(yhat.bag==df_test$default.payment.next.month)
```

```{r}
importance(bag.default)
varImpPlot(bag.default)
```

```{r}
yhat.bag = predict(bag.default,newdata=df_test,type="prob")
pred <- prediction(yhat.bag[,2], df_test$default.payment.next.month)
perf <- performance(pred,"tpr","fpr")

# plotting the ROC curve
plot(perf,col="black",lty=3, lwd=3)
auc <- performance(pred, measure = "auc")
auc <- auc@y.values[[1]]
auc
# KS STATISTIC
ks <- max(perf@y.values[[1]]-perf@x.values[[1]])
ks

#library(MLmetrics)
#default_num=as.numeric(df_test$default.payment.next.month)
#Gini(y_pred = bag.default, y_true = default_num)
```


```{r message=FALSE, warning=FALSE}
# Grows the forest
set.seed(1)
fit.forest <- randomForest(default.payment.next.month~., data=df,subset=train_ind, na.action=na.roughfix, mtry=5,ntree=200, importance=TRUE)

yhat.bag = predict(fit.forest,newdata=df_test,type="class")
table(yhat.bag,df_test$default.payment.next.month)
mean(yhat.bag==df_test$default.payment.next.month)

# Determine variable importance

importance(fit.forest)
varImpPlot(fit.forest)
```

```{r}
yhat.bag = predict(fit.forest,newdata=df_test,type="prob")
pred <- prediction(yhat.bag[,2], df_test$default.payment.next.month)
perf <- performance(pred,"tpr","fpr")

# plotting the ROC curve
plot(perf,col="black",lty=3, lwd=3)
auc <- performance(pred, measure = "auc")
auc <- auc@y.values[[1]]
auc
# KS STATISTIC
ks <- max(perf@y.values[[1]]-perf@x.values[[1]])
ks

library(MLmetrics)
default_num=as.numeric(df_test$default.payment.next.month)
Gini(y_pred = fit.foresst, y_true = default_num)
```

```{r,warning=FALSE,message=FALSE}
hc.complete=hclust(dist(x), method="complete")
plot(hc.complete,main="Complete Linkage", cex=.9)
```

